Equivalence Testing and the Second Generation \emph{P}-Value

### 2019-03-10 manuscript.R1.tex
R1 final

Second generation \emph{p}-values (SGPV) were recently proposed to as a
statistic that represents \enquote{the proportion of data-supported
hypotheses that are also null hypotheses} (Blume et al., 2018). The
researcher specifies an equivalence range around a null hypothesis of
values that are considered practically equivalent to the null
hypothesis. The SGPV measures the degree to which a set of
data-supported parameter values falls within the interval null
hypothesis. If the estimation interval falls completely within the
equivalence range, the SGPV is 1. If the confidence interval falls
completely outside of the equivalence range, the SGPV is 0. Otherwise
the SGPV is a value between 0 and 1 that expresses the overlap of
data-supported hypotheses and the equivalence range. When calculating
the SGPV the set of data-supported parameter values can be represented
by a confidence interval (CI), although one could also choose to use
credible intervals or Likelihood support intervals (SI). When a
confidence interval is used, the SGPV and equivalence tests such as the
Two One-Sided Tests (TOST) procedure (Lakens, 2017; Meyners, 2012;
Quertemont, 2011; Schuirmann, 1987) appear to have close ties, because
both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the
TOST procedure and the SGPV. We limit our analysis to continuous data
sampled from a bivariate normal distribution.

@@

The second generation \emph{p}-value (SGPV) is calculated as: \[
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
\] where I is the interval based on the data (e.g., a 95\% confidence
interval) and H\textsubscript{0} is the equivalence range. The first
term of this formula implies that the second generation \emph{p}-value
is the width of the confidence interval that overlaps with the
equivalence range, divided by the total width of the confidence
interval. The second term is a \enquote{small sample correction} (which
will be discussed later) that comes into play whenever the confidence
interval is more than twice as wide as the equivalence range. To examine
the relation between the TOST \emph{p}-value and the SGPV we can
calculate both statistics across a range of observed effect sizes. In
Figure \ref{fig:TOSTSGPV1} \emph{p}-values are plotted for the TOST
procedure and the SGPV. The statistics are calculated for hypothetical
one-sample \emph{t}-tests for observed means ranging from 140 to 150 (on
the x-axis). The equivalence range is set to 145 ± 2 (i.e., an
equivalence range from 143 to 147), the observed standard deviation is
assumed to be 2, and the sample size is 30. For example, for the
left-most point in Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST
\emph{p}-value is calculated for a hypothetical study with a sample size
of 30, an observed standard deviation of 2, and an observed mean of 140,
where the \emph{p}-value for the equivalence test is 1, and the SGPV is
0. Our conclusions about the relationship between TOST \emph{p}-values
and SGPV hold for an SGPV calculated from confidence intervals, and
assuming continuous normally distributed data. Readers can explore the
relationship between TOST \emph{p}-values and SGPV for themselves in an
online Shiny app: \url{http://shiny.ieis.tue.nl/TOST_vs_SGPV/}.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV1-1.pdf}
\caption{\label{fig:TOSTSGPV1}Comparison of \emph{p}-values from TOST (black
line) and SGPV (dotted grey line) across a range of observed sample
means (x-axis) tested against a mean of 145 in a one-sample
\emph{t}-test with a sample size of 30 and a standard deviation of 2.}
\end{figure}

The SGPV treats the equivalence range as the null-hypothesis, while the
TOST procedure treats the values outside of the equivalence range as the
null-hypothesis. For ease of comparison we can plot 1-SGPV (see Figure
\ref{fig:TOSTSGPV2}) to make the values more easily comparable. We see
that the \emph{p}-value from the TOST procedure and the SGPV follow each
other closely. When we discuss the relationship between the p-values
from TOST and the SGPV, we focus on their correspondence at three
values, namely where the TOST \emph{p} = 0.025 and SGPV is 1, where the
TOST \emph{p} = 0.5 and SGPV = 0.5, and where the TOST \emph{p} = 0.975
and SGPV = 1. These three values are important for the SGPV because they
indicate the values at which the SGPV indicates the data should be
interpreted as compatible with the null hypothesis (SGPV = 1), or with
the alternative hypothesis (SGPV = 0), or when the data are strictly
inconclusive (SGPV = 0.5). These three points of overlap are indicated
by the horizontal dotted lines in Figure \ref{fig:TOSTSGPV2}.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV2-1.pdf}
\caption{\label{fig:TOSTSGPV2}Comparison of \emph{p}-values from TOST (black
line) and 1-SGPV (dotted grey line) across a range of observed sample
means (x-axis) tested against a mean of 145 in a one-sample
\emph{t}-test with a sample size of 30 and a standard deviation of 2.}
\end{figure}

@@

Two other points always have to overlap. When the 95\% CI falls
completely inside the equivalence region, and one endpoint of the
confidence interval is exactly equal to one of the equivalence bounds
(see situation B in Figure \ref{fig:TOSTSGPV3}) the TOST \emph{p}-value
(which relies on a one-sided test) is always 0.025, and the SGPV is 1.
The third point where the SGPV and the \emph{p}-value from the TOST
procedure should overlap is where the 95\% CI falls completely outside
of the equivalence range, but one endpoint of the confidence interval is
equal to the equivalence bound (see situation C in Figure
\ref{fig:TOSTSGPV3}), when the \emph{p}-value will always be 0.975, and
the SGPV is 0. Note that this situation is in essence a minimum-effect
test (Murphy, Myors, \& Wolach, 2014). The goal of a minimum-effect is
not just to reject a difference of zero, but to reject the smallest
effect size of interest (i.e., the equivalence bounds). An equivalence
test and minimum effect test against the same equivalence bound are
complementary, and when a TOST \emph{p}-value is larger than 0.975, the
\emph{p}-value for the minimum effect test is smaller than 0.05, and the
minimum effect test provides no additional information that can not be
derived from the \emph{p}-value from the equivalence test. The SGPV
summarizes the information from an equivalence test and a minimum-effect
test. These can be two relevant questions to ask, although it often
makes sense to combine an equivalence test and a null-hypothesis test
instead (Lakens et al., 2018).

@@

For example, in Figure \ref{fig:TOSTSGPV4} we have plotted four SGPV's.
From A to D the SGPV is 0.76, 0.81, 0.86, and 0.91. The difference in
the percentage of overlap between A and B (-0.05) is identical to the
difference in the percentage of overlap between C and D as the mean gets
0.1 closer to the test value (-0.05). As the observed mean in a
one-sample \emph{t}-test lies closer to the test value, from situation A
to D, the mean gets closer to the test value by 0.1) the difference in
the overlap changes uniformly. As we move the observed mean closer to
the test value in steps of 0.1 across A to D the \emph{p}-value
calculated for normally distributed data are not uniformly distributed.
The probability of observing data more extreme than the upper bound of 2
is (from A to D) 0.16, 0.12, 0.08, and 0.06. As we can see, the
difference between A and B (0.04) is not the same as the difference
between C and D (0.03). Indeed, the difference in \emph{p}-values is the
largest as you start at \emph{p} = 0.5 (when the observed mean falls on
the test value), which is why the line in Figure \ref{fig:TOSTSGPV1} is
the steepest at \emph{p} = 0.5. Note that where the SGPV reaches 1 or 0,
\emph{p}-values closely approximate 0 and 1, but never reach these
values.

@@

There are three situations where \emph{p}-values for TOST differentiate
between observed results, while the SGPV does not differentiate. The
first two situations were discussed before and can be seen in Figure
\ref{fig:TOSTSGPV1}. When the SGPV is either 0 or 1, \emph{p}-values
from the equivalence test fall between 0.975 and 1 or between 0 and
0.025. While the SGPV is 1 as long as the confidence interval falls
completely within the equivalence bounds, the \emph{p}-value for the
TOST continues to differentiate between results as a function of how far
the confidence interval lies within the equivalence bounds (the further
the confidence interval is from both bounds, the lower the
\emph{p}-value). The easiest way to see this is by plotting the SGPV
against the \emph{p}-value from the TOST procedure. The situations where
the \emph{p}-values from the TOST procedure continue to differentiate
based on how extreme the results are, but the SGPV is a fixed value are
indicated by the parts of the curve where there are vertical straight
lines at second generation \emph{p}-values of 0 and 1.

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV6-1.pdf}
\caption{\label{fig:TOSTSGPV6}Comparison of \emph{p}-values from TOST (black
line) and SGPV (dotted grey line) across a range of observed sample
means (x-axis). Because the sample size is small (n = 10) and the CI is
more than twice as wide as the equivalence range (set to -0.4 to 0.4),
the SGPV is set to 0.5 (horizontal lightgrey line) across a range of
observed means.}
\end{figure}

We can examine this situation by calculating the SGPV and performing the
TOST for a situation where sample sizes are small and the equivalence
range is narrow, such that the CI is more than twice as large as the
equivalence range (see Figure \ref{fig:TOSTSGPV6}). When the two
statistics are plotted against each other we can see where the SGPV is
the same while the TOST \emph{p}-value still differentiates different
observed means (indicated by straight lines in the curve, see Figure
\ref{fig:TOSTSGPV7}). We see the SGPV is 0.5 for a range of observed
means where the \emph{p}-value from the equivalence test still varies.
It should be noted that in these calculations the \emph{p}-values for
the TOST procedure are \emph{never} smaller than 0.05 (i.e., they do not
get below 0.05 on the y-axis). In other words, we cannot conclude
equivalence based on any of the observed means. This happens because the
TOST \emph{p}-value is smaller than 0.05 only when the 90\% CI falls
completely between the upper and lower equivalence bounds. However, we
are examining a scenario where the 90\% CI is so wide that it never
falls completely within the two equivalence bounds, and thus the
equivalence test is never significant. As Lakens (2017) notes:
\enquote{in small samples (where CIs are wide), a study might have no
statistical power (i.e., the CI will always be so wide that it is
necessarily wider than the equivalence bounds).} None of the
\emph{p}-values based on the TOST procedure are below 0.05, and thus, in
the long run we have 0\% power.

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV9-1.pdf}
\caption{\label{fig:TOSTSGPV9}Comparison of \emph{p}-values from TOST (black
line) and SGPV (dotted grey line) across a range of observed sample
means (x-axis). The sample size is small (n = 10), but because the sd is
half as big as in Figure \ref{fig:TOSTSGPV7} (1 instead of 2) the CI is
less than twice as wide as the equivalence range (set to -0.4 to 0.4).
The SGPV is not set to 0.5 (horizontal light grey line) but reaches a
maximum slightly above 0.5 across a range of observed means.}
\end{figure}

@@

So far we have only looked at the relation between equivalence tests and
the SGPV when confidence intervals are symmetric (e.g., for confidence
intervals around mean differences). For correlations, which are bound
between -1 and 1, confidence intervals are only symmetric for a
correlation of exactly 0. The confidence interval for a correlation
becomes increasingly asymmetric as the observed correlation nears -1 or
1. For example, with ten observations, an observed correlation of 0 has
a symmetric 95\% confidence interval ranging from -0.630 to 0.630, while
and observed correlation of 0.7 has an asymmetric 95\% confidence
interval ranging from 0.126 to 0.993. Note that calculating confidence
intervals for a correlation involves a Fisher's z-transformation, which
transforms values such that they are approximately normally distributed,
which allows one to compute symmetric confidence intervals, which are
then retransformed into a correlation, where the confidence intervals
are asymmetric if the correlation is not exactly zero.

The effect of asymmetric confidence intervals around correlations is
most noticeable at smaller sample sizes. In Figure \ref{fig:TOSTSGPV11}
we plot the \emph{p}-values from equivalence tests and the SGPV (again
plotted as 1-SGPV for ease of comparison) for correlations. The sample
size is 30 pairs of observations, and the lower and upper equivalence
bounds are set to -0.45 and 0.45, with an alpha of 0.05. As the observed
correlation in the sample moves from -1 to 0 the \emph{p}-value from the
equivalence test becomes smaller, as does 1-SGPV. The pattern is quite
similar to that in Figure \ref{fig:TOSTSGPV2}. The \emph{p}-value for
the TOST procedure and 1-SGPV are still identical when \emph{p}-values
are 0.975 and 0.025 (indicated by the upper and lower horizontal dotted
lines). There are two important differences, however. First of all, the
SGPV is no longer a straight line, but a curve, due to the asymmetry in
the 95\% CI. Second, and most importantly, the \emph{p}-value for the
equivalence test and the SGPV do no longer overlap at \emph{p} = 0.5.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV11-1.pdf}
\caption{\label{fig:TOSTSGPV11}Comparison of \emph{p}-values from TOST
(black line) and 1-SGPV (dotted grey curve) across a range of observed
sample correlations (x-axis) tested against equivalence bounds of r =
-0.45 and r = 0.45 with n = 30 and an alpha of 0.05.}
\end{figure}

The reason that the equivalence test and SGPV no longer overlap is also
because of asymmetric confidence intervals. If the observed correlation
falls exactly on the equivalence bound the \emph{p}-value for the
equivalence test indicates that the probability of observing the
observed or more extreme data, assuming the equivalence bound is the
true effect size, is 50\%. In other words, if the true effect size is
the same as the equivalence bound, it is equally likely to find an
effect more extreme than the equivalence bound, as it is to observe an
effect that is less extreme than the equivalence bound. However, as can
be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation
\emph{p}-values associated with the observed correlations at r = -0.45
and r = 0.45 are larger than 50\%. Because the confidence intervals are
asymmetric around the observed effect size of 0.45 (ranging from 0.11 to
0.70) according to Blume et al. (2018) 58.11\% of the data-supported
hypotheses are null hypotheses, and therefore 58.10\% of the
data-supported hypotheses are compatible with the null premise.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV12-1.pdf}
\caption{\label{fig:TOSTSGPV12}Three 95\% confidence intervals for observed
effect sizes of r = -0.45, r = 0, and r = 0.45 for n = 30. Only the
confidence interval for r = 0 is symmetric.}
\end{figure}

@@

When introducing a new statistical method, it is important to compare it
to existing approaches and specify its relative strengths and
weaknesses. First of all, even though a SGPV of 1 or 0 has a clear
interpretation (we can reject effects outside or inside the equivalence
range), intermediate values are not as easy to interpret (especially for
effects that have asymmetric confidence intervals). In one sense, they
are what they are (the proportion of overlap), but it can be unclear
what this number tells us about the data we have collected. This is not
too problematic, since the main use of the SGPV (e.g., in all examples
provided by Blume and colleagues) seems to be to examine whether the
SGPV is 0, 1, or inconclusive. As already mentioned, this interpretation
of a SGPV is very similar to the Neyman-Pearson interpretation of an
equivalence test and a minimum effect tests (which are complementary).
The difference is that where a SGPV of 1 can be interpreted as \emph{p}
\textless{} .025, equivalence tests provide exact \emph{p}-values, and
they continue to differentiate between for example \emph{p} = 0.048 and
\emph{p} = 0.002. Whether this is desirable depends on the perspective
that is used. From a Neyman-Pearson perspective on statistical
inferences the main conclusion is based on whether or not
\(p < \alpha\), and thus an equivalence test and SGPV can be performed
by simply checking whether the confidence interval falls within the
equivalence range, just as a null-hypothesis test can be performed by
checking whether the confidence interval contains zero or not. At the
same time, it is recommended to report exact \emph{p}-values (American
Psychological Association, 2010), and exact \emph{p}-values might
provide information of interest to readers about how precisely how
surprising the data, or more extreme data, is under the null model. Some
researchers might be interested in combining an equivalence tests with a
null-hypothesis significance test. This allows a researcher to ask
whether there is an effect that is statistically different from zero,
and whether effect sizes that are considered meaningful can be rejected.
Equivalence tests combined with null-hypothesis tests classify results
into four possible categories, and for example allow researchers to
conclude an effect is significant \emph{and} equivalent (i.e.,
statistically different from zero, but also too small to be considered
meaningful, see Lakens et al. (2018)).

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV13-1.pdf}
\caption{\label{fig:TOSTSGPV13}Comparison of \emph{p}-values from TOST
(black line) and 1-SGPV (dotted grey curve) across a range of observed
sample correlations (x-axis) tested against equivalence bounds of r =
0.4 and r = 0.8 with n = 10 and an alpha of 0.05.}
\end{figure}

@@

Blume et al. (2018) claim that \enquote{Adjustments for multiple
comparisons are obviated} (p.~15). However, this is not correct. Given
the direct relationship between TOST and SGPV highlighted in this
manuscript (where to TOST \emph{p} = 0.025 equals SGPV = 1, as long as
the SGPV is calculated based on confidence intervals, and assuming data
sampled from a continuous bivariate normal distribution), not correcting
for multiple comparisons will inflate the probability of concluding the
absence of a meaningful effect based on the SGPV in exactly the same way
as for equivalence tests. Whenever statistical tests are interpreted as
support for a hypothesis (e.g., SPGV = 0 or SGPV =1), it is possible do
so erroneously, and if researchers want to control error rates, they
need to correct for multiple comparisons.

>>>>

Second generation \emph{p}-values (SGPV) were recently proposed as a
statistic that represents \enquote{the proportion of data-supported
hypotheses that are also null hypotheses} (Blume et al., 2018). The
researcher specifies an equivalence range around a null hypothesis of
values that are considered practically equivalent to the null
hypothesis. The SGPV measures the degree to which a set of
data-supported parameter values falls within the interval null
hypothesis. If the estimation interval falls completely within the
equivalence range, the SGPV is 1. If the confidence interval falls
completely outside of the equivalence range, the SGPV is 0. Otherwise
the SGPV is a value between 0 and 1 that expresses the overlap of
data-supported hypotheses and the equivalence range. When calculating
the SGPV the set of data-supported parameter values can be represented
by a confidence interval (CI), although one could also choose to use
credible intervals or Likelihood support intervals (SI). When a
confidence interval is used, the SGPV and equivalence tests such as the
Two One-Sided Tests (TOST) procedure (Lakens, 2017; Meyners, 2012;
Quertemont, 2011; Schuirmann, 1987) appear to have close ties, because
both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the
TOST procedure and the SGPV. We limit our analysis to continuous data
sampled from a bivariate normal distribution.

@@

The second generation \emph{p}-value (SGPV) is calculated as: \[
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
\] where I is the interval based on the data (e.g., a 95\% confidence
interval) and H\textsubscript{0} is the equivalence range. The first
term of this formula implies that the second generation \emph{p}-value
is the width of the confidence interval that overlaps with the
equivalence range, divided by the total width of the confidence
interval. The second term is a \enquote{small sample correction} (which
will be discussed later) that comes into play whenever the confidence
interval is more than twice as wide as the equivalence range. To examine
the relation between the TOST \emph{p}-value and the SGPV we can
calculate both statistics across a range of observed effect sizes.
Replicating the example by Blume et al. (2018), in Figure
\ref{fig:TOSTSGPV1} \emph{p}-values are plotted for the TOST procedure
and the SGPV. The statistics are calculated for hypothetical one-sample
\emph{t}-tests for observed means ranging from 140 to 150 (on the
x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence
range from 143 to 147), the observed standard deviation is assumed to be
2, and the sample size is 30. For example, for the left-most point in
Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST \emph{p}-value is
calculated for a hypothetical study with a sample size of 30, an
observed standard deviation of 2, and an observed mean of 140, where the
\emph{p}-value for the equivalence test is 1, and the SGPV is 0.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV1-1.pdf}
\caption{\label{fig:TOSTSGPV1}Comparison of \emph{p}-values from TOST (black
line) and SGPV (grey line) across a range of observed sample means
(x-axis) tested against a mean of 145 in a one-sample \emph{t}-test with
a sample size of 30 and a standard deviation of 2, illustrating that
when the TOST \emph{p}-value = 0.5, the SGPV = 0.5, when the TOST
\emph{p}-value is 0.975, 1-SGPV = 1, and when the TOST \emph{p}-value =
0.025, 1-SGPV = 0.}
\end{figure}

Our conclusions about the relationship between TOST \emph{p}-values and
SGPV hold for second generation \emph{p}-values calculated from
confidence intervals, and assuming data is sampled from a bivariate
normal distribution. Readers can explore the relationship between TOST
\emph{p}-values and SGPV for themselves in an online Shiny app:
\url{http://shiny.ieis.tue.nl/TOST_vs_SGPV/}.

The SGPV treats the equivalence range as the null-hypothesis, while the
TOST procedure treats the values outside of the equivalence range as the
null-hypothesis. For ease of comparison we can plot 1-SGPV (see Figure
\ref{fig:TOSTSGPV2}) to make the values more easily comparable. We see
that the \emph{p}-value from the TOST procedure and the SGPV follow each
other closely. When we discuss the relationship between the
\emph{p}-values from TOST and the SGPV, we focus on their correspondence
at three values, namely where the TOST \emph{p} = 0.025 and SGPV is 1,
where the TOST \emph{p} = 0.5 and SGPV = 0.5, and where the TOST
\emph{p} = 0.975 and SGPV = 1. These three values are important for the
SGPV because they indicate the values at which the SGPV indicates the
data should be interpreted as compatible with the null hypothesis (SGPV
= 1), or with the alternative hypothesis (SGPV = 0), or when the data
are strictly inconclusive (SGPV = 0.5). These three points of overlap
are indicated by the horizontal dotted lines in Figure
\ref{fig:TOSTSGPV2} at TOST \emph{p}-values of 0.975, 0.5, and 0.025.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV2-1.pdf}
\caption{\label{fig:TOSTSGPV2}Comparison of \emph{p}-values from TOST (black
line) and 1-SGPV (grey line) across a range of observed sample means
(x-axis) tested against a mean of 145 in a one-sample \emph{t}-test with
a sample size of 30 and a standard deviation of 2.}
\end{figure}

@@

Two other points always have to overlap. When the 95\% CI falls
completely inside the equivalence region, and one endpoint of the
confidence interval is exactly equal to one of the equivalence bounds
(see situation B in Figure \ref{fig:TOSTSGPV3}) the TOST \emph{p}-value
(which relies on a one-sided test) is always 0.025, and the SGPV is 1.
Note that when sample sizes are small or equivalence bounds are narrow,
small \emph{p}-values for the TOST or a SGPV = 1 might not be observed
in practice if too few observations are collected. The third point where
the SGPV and the \emph{p}-value from the TOST procedure should overlap
is where the 95\% CI falls completely outside of the equivalence range,
but one endpoint of the confidence interval is equal to the equivalence
bound (see situation C in Figure \ref{fig:TOSTSGPV3}), when the
\emph{p}-value will always be 0.975, and the SGPV is 0. Note that this
situation is in essence a minimum-effect test (Murphy, Myors, \& Wolach,
2014). The goal of a minimum-effect is not just to reject a difference
of zero, but to reject the smallest effect size of interest (i.e., the
equivalence bounds). An equivalence test and minimum effect test against
the same equivalence bound are complementary, and when a TOST
\emph{p}-value is larger than 0.975, the \emph{p}-value for the minimum
effect test is smaller than 0.05 (and therefore the minimum effect test
provides no additional information that can not be derived from the
\emph{p}-value from the equivalence test). The SGPV summarizes the
information from an equivalence test (and the complementary
minimum-effect test). These can be two relevant questions to ask,
although it often makes sense to combine an equivalence test and a
null-hypothesis test instead (Lakens et al., 2018).

@@

For example, in Figure \ref{fig:TOSTSGPV4} we have plotted four SGPV's.
From A to D the SGPV is 0.76, 0.81, 0.86, and 0.91. The difference in
the percentage of overlap between A and B (-0.05) is identical to the
difference in the percentage of overlap between C and D as the mean gets
0.1 closer to the test value (-0.05). As the observed mean in a
one-sample \emph{t}-test lies closer to the test value, from situation A
to D, the difference in the overlap changes uniformly. As we move the
observed mean closer to the test value in steps of 0.1 across A to D the
\emph{p}-value calculated for normally distributed data are not
uniformly distributed. The probability of observing data more extreme
than the upper bound of 2 is (from A to D) 0.16, 0.12, 0.08, and 0.05.
As we can see, the difference between A and B (0.04) is not the same as
the difference between C and D (0.03). Indeed, the difference in
\emph{p}-values is the largest as you start at \emph{p} = 0.5 (when the
observed mean falls on the test value), which is why the line in Figure
\ref{fig:TOSTSGPV1} is the steepest at \emph{p} = 0.5. Note that where
the SGPV reaches 1 or 0, \emph{p}-values closely approximate 0 and 1,
but never reach these values.

@@

There are three situations where \emph{p}-values for TOST differentiate
between observed results, while the SGPV does not differentiate. The
first two situations were discussed before and can be seen in Figure
\ref{fig:TOSTSGPV1}. When the SGPV is either 0 or 1, \emph{p}-values
from the equivalence test fall between 0.975 and 1 or between 0 and
0.025. Where the SGPV is 1 as long as the confidence interval falls
completely within the equivalence bounds, the \emph{p}-value for the
TOST continues to differentiate between results as a function of how far
the confidence interval lies within the equivalence bounds (the further
the confidence interval is from both bounds, the lower the
\emph{p}-value). The easiest way to see this is by plotting the SGPV
against the \emph{p}-value from the TOST procedure. The situations where
the \emph{p}-values from the TOST procedure continue to differentiate
based on how extreme the results are, but the SGPV is a fixed value are
indicated by the parts of the curve where there are vertical straight
lines at second generation \emph{p}-values of 0 and 1.

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV6-1.pdf}
\caption{\label{fig:TOSTSGPV6}Comparison of \emph{p}-values from TOST (black
line) and SGPV (grey line) across a range of observed sample means
(x-axis). Because the sample size is small (n = 10) and the CI is more
than twice as wide as the equivalence range (set to -0.4 to 0.4), the
SGPV is set to 0.5 (horizontal lightgrey line) across a range of
observed means.}
\end{figure}

We can examine this situation by calculating the SGPV and performing the
TOST for a situation where sample sizes are small and the equivalence
range is narrow, such that the CI is more than twice as large as the
equivalence range (see Figure \ref{fig:TOSTSGPV6}). When the two
statistics are plotted against each other we can see where the SGPV is
the same while the TOST \emph{p}-value still differentiates different
observed means (indicated by straight lines in the curve, see Figure
\ref{fig:TOSTSGPV7}). We see the SGPV is 0.5 for a range of observed
means where the \emph{p}-value from the equivalence test still varies.
It should be noted that in these calculations the \emph{p}-values for
the TOST procedure are \emph{never} smaller than 0.05 (i.e., they do not
get below 0.05 on the y-axis). In other words, we cannot conclude
equivalence based on any of the observed means. This happens because the
we are examining a scenario where the 90\% CI is so wide that it never
falls completely within the two equivalence bounds. As Lakens (2017)
notes: \enquote{in small samples (where CIs are wide), a study might
have no statistical power (i.e., the CI will always be so wide that it
is necessarily wider than the equivalence bounds).} None of the
\emph{p}-values based on the TOST procedure are below 0.05, and thus, in
the long run we have 0\% power.

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV9-1.pdf}
\caption{\label{fig:TOSTSGPV9}Comparison of \emph{p}-values from TOST (black
line) and SGPV (grey line) across a range of observed sample means
(x-axis). The sample size is small (n = 10), but because the sd is half
as big as in Figure \ref{fig:TOSTSGPV7} (1 instead of 2) the CI is less
than twice as wide as the equivalence range (set to -0.4 to 0.4). The
SGPV is not set to 0.5 (horizontal light grey line) but reaches a
maximum slightly above 0.5 across a range of observed means.}
\end{figure}

@@

So far we have only looked at the relation between equivalence tests and
the SGPV when confidence intervals are symmetric (e.g., for confidence
intervals around mean differences). For correlations, which are bound
between -1 and 1, confidence intervals are only symmetric for a
correlation of exactly 0. The confidence interval for a correlation
becomes increasingly asymmetric as the observed correlation nears -1 or
1. For example, with ten observations, an observed correlation of 0 has
a symmetric 95\% confidence interval ranging from -0.630 to 0.630, while
and observed correlation of 0.7 has an asymmetric 95\% confidence
interval ranging from 0.126 to 0.993. Note that calculating confidence
intervals for a correlation involves a Fisher's z-transformation, which
transforms values such that they are approximately normally
z-distributed, which allows one to compute symmetric confidence
intervals. These confidence intervals are then retransformed into a
correlation, where the confidence intervals are asymmetric if the
correlation is not exactly zero.

The effect of asymmetric confidence intervals around correlations is
most noticeable at smaller sample sizes. In Figure \ref{fig:TOSTSGPV11}
we plot the \emph{p}-values from equivalence tests and the SGPV (again
plotted as 1-SGPV for ease of comparison) for correlations. The sample
size is 30 pairs of observations, and the lower and upper equivalence
bounds are set to -0.45 and 0.45, with an alpha of 0.05. As the observed
correlation in the sample moves from -1 to 0 the \emph{p}-value from the
equivalence test becomes smaller, as does 1-SGPV. The pattern is quite
similar to that in Figure \ref{fig:TOSTSGPV2}. The \emph{p}-value for
the TOST procedure and 1-SGPV are still related as discussed above, with
TOST \emph{p}-values of 0.975 and 0.025 corresponding to a 1-SGPV of 1
and 0, respectively. There are two important differences, however. First
of all, the SGPV is no longer a straight line, but a curve, due to the
asymmetry in the 95\% CI. Second, and most importantly, the
\emph{p}-value for the equivalence test and the SGPV do no longer
overlap at \emph{p} = 0.5.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV11-1.pdf}
\caption{\label{fig:TOSTSGPV11}Comparison of \emph{p}-values from TOST
(black line) and 1-SGPV (grey curve) across a range of observed sample
correlations (x-axis) tested against equivalence bounds of r = -0.45 and
r = 0.45 with n = 30 and an alpha of 0.05.}
\end{figure}

The reason that the equivalence test and SGPV no longer overlap is also
because of asymmetric confidence intervals. If the observed correlation
falls exactly on the equivalence bound the \emph{p}-value for the
equivalence test indicates that the probability of observing the
observed or more extreme data, assuming the equivalence bound is the
true effect size, is 50\%. In other words, if the true effect size is
the same as the equivalence bound, it is equally likely to find an
effect more extreme than the equivalence bound, as it is to observe an
effect that is less extreme than the equivalence bound. However, as can
be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation
\emph{p}-values associated with the observed correlations at \emph{r} =
-0.45 and \emph{r} = 0.45 are larger than 50\%. Because the confidence
intervals are asymmetric around the observed effect size of 0.45
(ranging from 0.11 to 0.70) according to Blume et al. (2018) 58.11\% of
the data-supported hypotheses are null hypotheses, and therefore 58.10\%
of the data-supported hypotheses are compatible with the null premise.

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV12-1.pdf}
\caption{\label{fig:TOSTSGPV12}Three 95\% confidence intervals for observed
effect sizes of \emph{r} = -0.45, \emph{r} = 0, and \emph{r} = 0.45 for
n = 30. Only the confidence interval for \emph{r} = 0 is symmetric.}
\end{figure}

@@

When introducing a new statistical method, it is important to compare it
to existing approaches and specify its relative strengths and
weaknesses. Here, we aimed to compare the SGPV against equivalence tests
based on the TOST procedure. First of all, even though a SGPV of 1 or 0
has a clear interpretation (we can reject effects outside or inside the
equivalence range), intermediate values are not as easy to interpret
(especially for effects that have asymmetric confidence intervals). In
one sense, they are what they are (the proportion of overlap), but it
can be unclear what this number tells us about the data we have
collected. This is not too problematic, since the main use of the SGPV
(e.g., in all examples provided by Blume and colleagues) seems to be to
examine whether the SGPV is 0, 1, or inconclusive. As already mentioned,
this interpretation of a SGPV is very similar to the Neyman-Pearson
interpretation of an equivalence test and a minimum effect tests (which
are complementary). The difference is that where a SGPV of 1 can be
interpreted as \emph{p} \textless{} .025, equivalence tests provide
exact \emph{p}-values, and they continue to differentiate between for
example \emph{p} = 0.024 and \emph{p} = 0.002. Whether this is desirable
depends on the perspective that is used. From a Neyman-Pearson
perspective on statistical inferences the main conclusion is based on
whether or not \(p < \alpha\), and thus an equivalence test and SGPV can
be performed by simply checking whether the confidence interval falls
within the equivalence range, just as a null-hypothesis test can be
performed by checking whether the confidence interval contains zero or
not. At the same time, it is recommended to report exact \emph{p}-values
(American Psychological Association, 2010), and exact \emph{p}-values
might provide information of interest to readers about how precisely how
surprising the data, or more extreme data, is under the null model. Some
researchers might be interested in combining an equivalence tests with a
null-hypothesis significance test. This allows a researcher to ask
whether there is an effect that is statistically different from zero,
and whether effect sizes that are considered meaningful can be rejected.
Equivalence tests combined with null-hypothesis tests classify results
into four possible categories, and for example allow researchers to
conclude an effect is significant \emph{and} equivalent (i.e.,
statistically different from zero, but also too small to be considered
meaningful, see Lakens et al. (2018)).

@@

\begin{figure}
\centering
\includegraphics{manuscript.R1_files/figure-latex/TOSTSGPV13-1.pdf}
\caption{\label{fig:TOSTSGPV13}Comparison of \emph{p}-values from TOST
(black line) and 1-SGPV (grey curve) across a range of observed sample
correlations (x-axis) tested against equivalence bounds of \emph{r} =
0.4 and \emph{r} = 0.8 with n = 10 and an alpha of 0.05.}
\end{figure}

@@

Blume et al. (2018) claim that when using the SGPV \enquote{Adjustments
for multiple comparisons are obviated} (p.~15). However, this is not
correct. Given the direct relationship between TOST and SGPV highlighted
in this manuscript (where a TOST \emph{p} = 0.025 equals SGPV = 1, as
long as the SGPV is calculated based on confidence intervals, and
assuming data are sampled from a continuous bivariate normal
distribution), not correcting for multiple comparisons will inflate the
probability of concluding the absence of a meaningful effect based on
the SGPV in exactly the same way as it will for equivalence tests.
Whenever statistical tests are interpreted as support for a hypothesis
(e.g., SPGV = 0 or SGPV = 1), it is possible do so erroneously, and if
researchers want to control error rates, they need to correct for
multiple comparisons.
